---
title: 'Week 6: Bilateral Trade'
author: "Xuan Pham"
date: "9/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, let's pull the export and import trade values between US and China from 1980 to 2018.  
```{r}
library(tradestatistics)
library(tidyverse)
library(xts)

options(scipen=999)
bilateral <- ots_create_tidy_data(year = c(1990:2018), reporters="us", partners="chn", table = "yrp")
```

We need to adjust the reported values for inflation. The latest inflation adjustment published by the World Bank is 2017. We have to convert all reported values to 2017 dollar amount.  

```{r}
bilateral.adj <- ots_inflation_adjustment(trade_data = bilateral,
reference_year=2017)
```

Using brute force to fill in the value for 2017.  

```{r}
year_2017 <- bilateral[28,"export_value_usd"]-bilateral[28,"import_value_usd"]
```

Now we can calculate the trade balance (Export - Import).  

```{r}
bilateral.df <- bilateral.adj %>% dplyr::mutate(balance = export_value_usd - import_value_usd) %>% select(year, balance)

bilateral.df[28,"balance"] <- year_2017
```

Transforming balance by dividing $1M. Balance is now reported in millions of dollars.  

```{r}
bilateral.df$balance <- bilateral.df$balance/1000000
```

Creating an xts object.  

```{r}
bilateral.df$mdy <- paste0('12-','31-',bilateral.df$year)
bilateral.df$mdy <- as.Date(bilateral.df$mdy,
                                  "%m-%d-%Y")
bilateral.xts <- xts(x=bilateral.df$balance,
                     order.by = bilateral.df$mdy)
plot(bilateral.xts)
```

Creating a ts object.  

```{r}
bilateral.ts <- ts(bilateral.df$balance, start=1990, end=2018, frequency=1)

autoplot(bilateral.ts)
```

Can you compare ARIMA and ETS models using cross validation?  
