---
title: 'Module 4: Smoothing, Part 2'
author: "Xuan Pham"
date: "9/10/2019"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simple Exponential Smoothing (SES)  

First, let's look at the data set for the week. Oil production in Saudi Arabia

```{r oildata}
library(forecast)
library(fpp2)

plot(oil) #entire series 

oildata <- window(oil, start = 1996, end = 2007) #subset of series
plot(oildata, ylab = "Oil (millions of tonnes)", xlab = "Year")
```

We know that we can deploy these simple forecasting methods:  
* mean 
* naive  
* moving average

What if we do not give equal weights to all observations? What if weights are assigned based on whether we think the more recent history or distant past is more important?  


```{r ses}
fit1 <- ses(oildata, alpha = 0.2, initial = "simple", h = 3)
fit2 <- ses(oildata, alpha = 0.6, initial = "simple", h = 3)
fit3 <- ses(oildata, h = 3) #note no value for alpha and no initial value.  The algorithm will calculate alpha as part of the optimization. 

#what kind of objects are these?
class(fit3)

#what is in them
summary(fit3)
summary(fit2)
```

Now let's look at some plots.   

```{r sesplots}
plot(fit1, PI=FALSE, ylab="Oil (millions of tonnes)",
     xlab="Year", main="", fcol="white", type="o")
lines(fitted(fit1), col="blue", type="o")
lines(fitted(fit2), col="red", type="o")
lines(fitted(fit3), col="green", type="o")

legend("topleft",lty=1, col=c(1,"blue","red","green"), cex = 0.75, 
       c("data", expression(alpha == 0.2), expression(alpha == 0.6),
         expression(alpha == 0.79)),pch=1)

```

What if we want to use our SES methods to forecast three years ahead? 2008-2010  

```{r forecasts}
fit1 <- ses(oildata, alpha = 0.2, initial = "simple", h = 3)
fit2 <- ses(oildata, alpha = 0.6, initial = "simple", h = 3)
fit3 <- ses(oildata, h = 3)

plot(fit1, PI=FALSE, ylim = c(400, 600), ylab="Oil (millions of tonnes)",
xlab="Year", main="", fcol="white", type="o")
#this just plots the data through 2007 because that's what we ran the model over
#we have more data but have to plot it seperately if we want to compare the forecast with the actual 
# add these lines for putting the actual data past 2007
test.oil <- window(oil, start = 2007)
lines(test.oil, col = "black", type = "o")
lines(fitted(fit1), col="blue", type="o")
lines(fitted(fit2), col="red", type="o")
lines(fitted(fit3), col="green", type="o")
lines(fit1$mean, col="blue", type="o")
lines(fit2$mean, col="red", type="o")
lines(fit3$mean, col="green", type="o")
legend("topleft",lty=1, cex =0.75, col=c(1,"blue","red","green"),
c("data", expression(alpha == 0.2), expression(alpha == 0.6),
expression(alpha == 0.79)),pch=1)
```

# Holt's Linear Trend  

The problem with SES is that it does not forecast anything else but a flat line. What if we expect forecasts to have a trend?  

```{r holt}
fith <- holt(oildata, alpha=0.8, beta=0.2, initial="simple", h=3) 
class(fith)
summary(fith)
oildata[1]
```

Notice that it shows the parameter settings we gave it. We now have a setting for the effect of the trend.  What's the initial value?  It's the first value of the oildata series - because we set the initialization routine = "simple"
fith$mean.  


Adding these lines to our chart.  What is the effect of adding trend? 
```{r holtplot}
plot(fith, PI=TRUE, ylim = c(400, 600))
lines(fitted(fith), type = "o", col="purple") 
lines(fith$mean, col="purple", type="o") 
lines(test.oil, col = "black", type = "o")
lines(fitted(fit1), col="blue", type="o")
lines(fitted(fit2), col="red", type="o")
lines(fitted(fit3), col="green", type="o")
lines(fit1$mean, col="blue", type="o")
lines(fit2$mean, col="red", type="o")
lines(fit3$mean, col="green", type="o")
legend("topleft",lty=1, cex = 0.75, col=c(1,"blue","red","green", "purple"),
       c("data", expression(alpha == 0.2), expression(alpha == 0.6),
         expression(alpha == 0.99),"Holt"),pch=1)
```

That was the additive trend Holt's method - let's make it muliplicative.  


```{r holtm}
fith2 <- holt(oildata,alpha=0.8,beta=0.2,initial="simple",exponential=TRUE,h=3) 
summary(fith2)
```
We specified alpha and beta and told it to start with the initial value and the initial trend.

Now add that line to the chart:
```{r hotmplot}
plot(fith2, PI=FALSE, ylim = c(400, 600))
lines(test.oil, col = "black", type = "o")
lines(fitted(fith2), col="brown", type = "o")
lines(fith2$mean, col="brown", type="o")
legend("topleft", cex =0.75, lty=1, col=c(1,"blue","red","green", "purple", "brown"),
       c("data", expression(alpha == 0.2), expression(alpha == 0.6),
         expression(alpha == 0.89),"Holt","Exponential"),pch=1)
```

Let's look closer at the components of each forecast output.

```{r forecastoutput}
fith2$model
fith2$x
fith2$fitted
fith2$mean
```

What about Holt's trend with a damped factor?  

```{r holtdamped}
fith3 <- holt(oildata,alpha=0.8,beta=0.2,initial="simple",exponential=TRUE,
              damped=TRUE, phi=NULL, h=3) #setting phi=NULL allows R to automatically come up with the best damped factor.
summary(fith3)

plot(fith3, PI=FALSE, ylim = c(400, 600))
lines(test.oil, col = "black", type = "o")
lines(fitted(fith3), col="brown", type = "o")
lines(fith3$mean, col="brown", type="o")
legend("topleft", cex =0.75, lty=1, col=c(1,"blue","red","green", "purple", "brown"),
       c("data", expression(alpha == 0.2), expression(alpha == 0.6),
         expression(alpha == 0.89),"Holt","Exponential"),pch=1)
```

#Holt-Winters' Method  

But what if there is seasonality in the time series?  

First, we need to find a different time series (with seasonality) to examine. Here's a time series of tourists coming to Australia.  

```{r}
#?austourists
plot(austourists, main = "International Tourists to Australia on Quarterly Basis (in Millions)")
aust <- window(austourists,start=2004, end=2014) #let's make the series a bit shorter for manageability purpose. 
fit1hw <- hw(aust,seasonal="additive")
fit2hw <- hw(aust,seasonal="multiplicative")

#Can you take a look at Holt Winters method with damped factor for trend?

test.aust <- window(austourists, start=2015)
plot(fit2hw,ylab="International visitor night in Australia (millions)",
     PI=FALSE, type="o", fcol="white", xlab="Year")
lines(test.aust, col = "black", type = "o")
lines(fitted(fit1hw), col="red", lty=2)
lines(fitted(fit2hw), col="green", lty=2)
lines(fit1hw$mean, type="o", col="red")
lines(fit2hw$mean, type="o", col="green")
legend("topleft",lty=1, cex = 0.75, pch=1, col=1:3, 
  c("data","Holt Winters' Additive","Holt Winters' Multiplicative"))
```
We can pull the different pieces out of the forecast object and compare them across specifications.  What's the difference between additive seasonal and multiplicative seasonal?

```{r}
states <- cbind(fit1hw$model$states[,1:3],fit2hw$model$states[,1:3])
colnames(states) <- c("level","slope","seasonal","level","slope","seasonal")
plot(states, xlab="Year", main = "Comparing Additive and Multiplicative Seasonals")
```

Next, let's try to pull out the seasonal indices.  


# ETS (Error-Trend-Seasonal) Statistical Models 

Let's revisit the oil data - we originally used the SES method on it which is equivalent to ETS(A,N,N)

```{r ets1}
oildata <- window(oil, start = 1996, end = 2007)
fita <- ets(oildata, model = "ANN")
plot(forecast(fita, h=3), ylab="Oil (millions of tonnes)")
summary(fita)
ls(fita)  #list names of the objects in the specified environment
fita$par
```

What if we used the ETS(M,N,N) model?

```{r ets2}
oildata <- window(oil, start = 1996, end = 2007)
fitm <- ets(oildata, model = "MNN")
plot(forecast(fitm, h=3), ylab="Oil (millions of tonnes)")
summary(fitm)
ls(fitm)  #list names of the objects in the specified environment
fitm$par

```
But are either of these the right model for these data?  What if we let the ETS code decide?

```{r ets3}
oildata <- window(oil, start = 1996, end = 2007)
fit <- ets(oildata)
summary(fit)
plot(fit)
plot(forecast(fit, h = 8), ylab = "Oil (millions of tonnes)")
```


The model chose additive.  Compare the AIC for the two we examined.

```{r aic}
fita$aic
fitm$aic
```
Yup - the AIC for the additive model is lower so it is the "better" model--but not by much.   

How about a more interesting example? Remember the Australian tourism time series? 

Very strongly seasonal.  Probably not a good candidate for the ETS(A,N,N) model.  Let's see:

```{r}
fit <- ets(aust)
summary(fit)
plot(fit)
plot(forecast(fit, h = 4), ylab = "Millions")
```

What about our ubiquitous air passengers data?

```{r}
ap <- window(AirPassengers)
fit <- ets(ap)
summary(fit)

plot(fit)

plot(forecast(fit, h = 8), ylab = "Millions")
```

