---
title: "Dissecting a Time Series"
output: html_document
---

# Time Series using ZOO

```{r}
library(tseries)
library(zoo)

#import dataset into a dataframe
cisco <- read.table('C:/Users/Xuan Pham/Dropbox/Fall_2019/BIA6315/code/Week 2/Data/cisco_00-10.csv', header=T, sep=',')

# create time series for cisco prices
ciscots <- zoo(cisco$Price, as.Date(as.character(cisco$Date), format = "%m/%d/%y"))


plot(ciscots)
hist(ciscots)
mean(ciscots)
```

#Creating new variables 

# Four moments of a normal distribution:
First moment: Mean 
Second moment: Variance
Third moment: Skewness
Fourth moment: Kurtosis

The goal is to determine whether the Cisco's daily rate of returns come from a normal distribution.
 
Looking at the histogram & time series plot, what do you see?

```{r} 
#Natural log transformation makes the distribution less skewed.
#DEFINE LOG RETURNS
#rts is a time series object since it is created from a TS object
log.ciscots <- log(ciscots)
hist(log.ciscots) #any difference?
plot(log.ciscots) #any difference?
mean(log.ciscots)
```

What about taking the day-to-day difference in log price? 

```{r }
#first the hard way of doing it...
lagged.ciscots <- lag(log.ciscots, k=-1)
price.diff <- (log.ciscots - lagged.ciscots)
head(price.diff)

#now the easier way...
rts = diff(log.ciscots, lag=1) #default is lag = 1
head(rts)
hist(rts) #any difference?
plot(rts) #any difference?
mean(rts)
```

This is called "de-trending" our time series. The mean is now close to 0. 


```{r}
rt <- coredata(rts) #keeping just the price data. No time index.
library(fBasics)
                     
# COMPUTE SUMMARY STATISTICS
basicStats(rt) 
           
# CREATE HISTOGRAM 
# OPTIONAL creates 2 by 2 display for 4 plots 
# par(mfcol=c(2,2)) 
hist(rt, xlab="Cisco log returns", prob=TRUE, main="Histogram") 
# add approximating normal density curve 
xfit<-seq(min(rt),max(rt),length=60) 
yfit<-dnorm(xfit,mean=mean(rt),sd=sd(rt)) 
lines(xfit, yfit, col="blue", lwd=2) 
```

Is there skewness?
What about kurtosis?

```{r}
# CREATE NORMAL PROBABILITY PLOT 
qqnorm(rt) 
qqline(rt, col = 2) 
plot(rt)
```


```{r}
library(moments)
skewness(rt)
kurtosis(rt)

# NORMALITY TESTS 
# Perform Jarque-Bera normality test. 
#H0: Data is normally distributed
#H1: Data is not normally distributed
normalTest(rt,method=c("jb")) 
#If you want to see the critical values for chi-square distribution, check here:
#https://www.itl.nist.gov/div898/handbook/eda/section3/eda3674.htm
```
So the Cisco price data set is not normally distributed. 

Now let's turn our attention to the "independent, identically distributed" assumption.

Independent: Each sample observation is uncorrelated with another sample observation.
Identically distributed: The sample observations are drawn from the same probability distribution.

It's obvious that time series violates this assumption.

How do we examine this?

In traditional statistics, we have covariance & correlation to examine the linear relationship between two variables, X & Y.

Covariance: Direction of the relationship
Correlation: Direction & strength of the relationship

In time series analysis, we change X & Y into Y(t) and Y(t-k) where k = some number of lagged periods. Hence, we have autocovariance and autocorrelation.

```{r}
# COMPUTE ACF AND PLOT CORRELOGRAM 
#prints acf values to console 
acf(rt, plot=F) 
#plot acf values on graph (correlogram) 
acf(rt, plot=T) 
```
Two things to look for in an ACF plot:
1. Do I have significant lags?
2. Do I have rapid/gradual decay?

Is there a hypothesis test we can run to test for autocorrelation? 

```{r}
# COMPUTE LJUNG-BOX TEST FOR WHITE NOISE (NO AUTOCORRELATION)
#H0: p(1) = p(2) = p(k) = 0
#H1: p(k) is not equal to 0

# to Lag 3
Box.test(rt,lag=3,type='Ljung')
# to Lag 13 
Box.test(rt,lag=13,type='Ljung') 
                     
```


